---
title: "Air Quality Index across year in NY state"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```


```{r, include=FALSE}
air_daily_df = 
  read_csv("./data/air_daily.csv") %>% 
  separate(col = date, into = c('year','month','day'), sep = "-" , convert = TRUE) %>% 
  drop_na(aqi) %>% 
  drop_na(county)
```


```{r, include=FALSE}
aqi_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, aqi) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    aqi_mean = mean(aqi)
  )
```

* Air quality index is different among counties in 10years. We created maps with their air quality index on colors among counties in NY state.If the color is grey, that means there was no respondents in those counties at that year.

#### Air quality index{.tabset} 

According to the graphs, we can know that:
* Overall, the air quality index is decreasing in 10 years in all counties, which means that the air quality is becoming better. 
* Urban counties, including `Bronx`, `Queens` , `New York` and `Erie`, are  with a higher air quality index, which means they have worse air quality.
* Based on the map, we can know that air quality index might with no relationship with asthma because in northern counties in NY state, where the asthma percentage are higher, the air quality are pretty good.

##### 2003

```{r}
air_county_df1 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2003) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map1 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df1, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map1
```

##### 2004

```{r}
air_county_df2 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2004) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map2 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df2, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map2
```

##### 2005

```{r}
air_county_df3 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2005) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map3 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df3, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map3
```

##### 2006

```{r}
air_county_df4 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2006) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map4 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df4, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map4
```

##### 2007

```{r}
air_county_df5 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2007) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map5 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df5, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map5
```

##### 2008

```{r}
air_county_df6 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2008) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map6 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df6, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map6
```

##### 2009

```{r}
air_county_df7 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2009) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map7 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df7, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map7
```

##### 2010

```{r}
air_county_df8 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2010) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map8 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df8, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map8
```

##### 2011

```{r}
air_county_df9 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2011) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map9 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df9, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map9
```

##### 2012

```{r}
air_county_df10 = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county, year) %>% 
  filter(year == 2012) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map10 = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df10, values = "aqi_all", labels = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "AQI", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map10
```
