---
title: "Analysis_in_website"
output: github_document
---
```{r setup, include=FALSE}
library(plyr)
library(tidyverse)
library(patchwork)
library(viridis)
library(haven)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```




# Data Import and Manipulation



  We first read in our cleaned data brfss_with_air2.csv and mutate the variable types as needed.

  We have asthma and asthma_now as the outcome variables, mean_co_month,mean_aqi_month, mental_health, physical_health, county, age, race, smoker and income as the predictors
```{r,message = FALSE}
asthma_df2 = 
  read_csv("./data/brfss_with_air2.csv") %>%
  mutate(
    #outcome
    asthma = as.numeric(asthma),
    asthma_now = as.numeric(asthma_now),
    #predictors below
    mean_aqi_month = as.numeric(mean_aqi_month),
    mental_health = as.numeric(mental_health),
    physical_health = as.numeric(physical_health),
    county = as.factor(county),
    age = as.factor(age),
    race = as.factor(race),
    smoker = as.factor(smoker),
    income = as.factor(income)
    )
  
```

# Linear Regression with Mean CO as Exposure


  
```{r,message = FALSE}
linear_fit_co_emergency = asthma_df2 %>%
  lm(asthma_emergency ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_co_month, data = .)

```

```{r}
linear_fit_co_emergency =
  linear_fit_co_emergency %>% 
  broom::tidy()
  
```

  We first ran a linear regression model with asthma_emergency as the outcome variable and check the correlation of mean_co_month with Asthma adjusting for other predictors.

```{r}
print(linear_fit_co_emergency, n=60)
```

**Interpretation**

  The results show that mean_co_month, the average CO concentration has a positive coefficient of 1.64 with Asthma_emergency occurrence. This means that higher CO concentration contributes to Asthma_emergency emergency with an 1.64 in cases per unit increase of mean_co_concentration.
  
  
  

  
# Logistic regression AQI as exposure

```{r}
fit_logistic1b = 
  asthma_df2 %>% 
  glm(asthma ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_aqi_month, data = ., family = binomial()) 
```
We then conducted a Logistic regression with asthma as the main outcome varibale and check the relationship between mean_aqi month with asthma status.


A logistic regression model was used because of the binary outcome variable.
```{r}
fit_logistic1b %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```

**Interpretation**

The results show that mean_aqi_month, the average Air quality index value, has an odd ratio of 1 towards Asthma status. This implies that mean_aqi_month does not have any influence on Asthma occurrences within our investigation.
