---
title: "Analysis_in_website"
output: github_document
---
```{r setup, include=FALSE}
library(plyr)
library(tidyverse)
library(patchwork)
library(viridis)
library(haven)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Data Import and Manipulation


* We first read in our cleaned data brfss_with_air2.csv and mutate the variable types as needed.

* We have asthma and asthma_now as the outcome variables, `mean_aqi_month`,`mean_so2_month`, `mean_no2_month`, `mean_co_month` and `mean_pm2_5_month` as predictors. `mental_health`, `physical_health`, `county`, `age`, `race`, `smoker` and `income` as the covariates.

```{r,message = FALSE}
asthma_df2 = 
  read_csv("./data/brfss_with_air2.csv") %>%
  mutate(
    #outcome
    asthma = as.numeric(asthma),
    asthma_now = as.numeric(asthma_now),
    #predictors below
    mean_aqi_month = as.numeric(mean_aqi_month),
    mental_health = as.numeric(mental_health),
    physical_health = as.numeric(physical_health),
    county = as.factor(county),
    age = as.factor(age),
    race = as.factor(race),
    smoker = as.factor(smoker),
    income = as.factor(income)
    )
  
```

# Linear Regression with asthma emergency visits time as outcome

* Based on our exploratory analysis, We first ran a linear regression model with `asthma_emergency` as the outcome variable and check the correlation with `mean_so2_month`, `mean_no2_month`, `mean_co_month` and `mean_pm2_5_month` adjusting for other predictors.

### SO2

```{r,message = FALSE}
linear_fit_so2_emergency = 
  asthma_df2 %>%
  lm(asthma_emergency ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_so2_month, data = .) %>% 
  broom::tidy() 

linear_fit_so2_emergency %>% 
  knitr::kable(digits = 3)
```

### NO2

```{r,message = FALSE}
linear_fit_no2_emergency = asthma_df2 %>%
  lm(asthma_emergency ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_no2_month, data = .) %>% 
  broom::tidy()

linear_fit_no2_emergency %>% 
  knitr::kable(digits = 3)
```

### CO

```{r,message = FALSE}
linear_fit_co_emergency = asthma_df2 %>%
  lm(asthma_emergency ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_co_month, data = .) %>% 
  broom::tidy()

linear_fit_co_emergency %>% 
  knitr::kable(digits = 3)
```


### PM2.5

```{r,message = FALSE}
linear_fit_pm2_5_emergency = asthma_df2 %>%
  lm(asthma_emergency ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_pm2_5_month, data = .)%>% 
  broom::tidy()

linear_fit_pm2_5_emergency %>% 
  knitr::kable(digits = 3)
```


**Interpretation**

* The results show that mean_so2_month, the average SO2 concentration has a positive coefficient of 0.166 with Asthma_emergency occurrence. This means that higher SO2 concentration contributes to Asthma emergency visit times with an 0.166 in cases per unit increase of average SO2 concentration in one month, adjusting for other covariate. The p-value is 0.000, which means that this is significant at a significance level of 0.05.

* The results show that mean_no2_month, the average NO2 concentration has a positive coefficient of 0.195 with Asthma_emergency occurrence. This means that higher NO2 concentration contributes to Asthma emergency visit times with an 0.195 in cases per unit increase of average NO2 concentration in one month, adjusting for other covariate. The p-value is 0.001, which means that this is significant at a significance level of 0.05.

* The results show that mean_co_month, the average CO concentration has a positive coefficient of 1.64 with Asthma_emergency occurrence. This means that higher CO concentration contributes to Asthma emergency visit times with an 1.64 in cases per unit increase of average CO concentration in one month, adjusting for other covariate. The p-value is 0.354, which means that this is not significant at a significance level of 0.05.

* The results show that mean_pm2_5_month, the average PM2.5 concentration has a positive coefficient of 0.237 with Asthma_emergency occurrence. This means that higher PM2.5 concentration contributes to Asthma emergency visit times with an 0.237 in cases per unit increase of average PM2.5 concentration in one month, adjusting for other covariate. The p-value is 0.287, which means that this is not significant at a significance level of 0.05.
  
  
  
# Logistic regression with asthma as exposure

* We still want to see if there is any relationship between `asthma` and `mean_aqi_month`. We then conducted a Logistic regression with `asthma` as the main outcome varibale and check the relationship between `mean_aqi_month` with asthma status. A logistic regression model was used because of the binary outcome variable.

```{r}
fit_logistic1b = 
  asthma_df2 %>% 
  glm(asthma ~ mental_health + physical_health + sex + smoker + race + age + income + county + mean_aqi_month, data = ., family = binomial()) 

fit_logistic1b %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```


**Interpretation**

The results show that mean_aqi_month, the average Air quality index value, has an odd ratio of 1 towards Asthma status. This implies that mean_aqi_month does not have any influence on Asthma occurrences within our investigation.
