---
title: "data_import_cleaning"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

check variables in daily parameters, and daily aqi tables:

```{r}
daily_ny_2022_co = read_csv("./data/daily_parameter/co/daily_42101_2022.csv") %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York")
```

```{r}
daily_ny_2022_no2 = read_csv("./data/daily_parameter/no2/daily_42602_2022.csv") %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York")
```

```{r}
daily_aqi = read_csv("./data/daily_aqi/daily_aqi_by_county_2022.csv")

daily_aqi_data = daily_aqi %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York")
```

## Daily aqi for each county in New York State for the last 15 years.

```{r}
full_aqi = 
  tibble(
    files = list.files("data/daily_aqi/"),
    path = str_c("data/daily_aqi/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(-files, -path) %>% 
  select(state = state_name, county = county_name, date, aqi, category, defining_parameter, defining_site, site_num = number_of_sites_reporting)
```

## daily parameter value

manipulate a relatively small file firstly:

```{r}
sample_ozone = read_csv("./data/daily_parameter/ozone/daily_44201_2019.csv") %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

There is a large file storage complication which would block the push. So we need to make a pre-processing for `ozone`, `pm2.5`, and `so2` folder (those folders contain raw data files that exceed 100 M). This process is in another document.

After the pre-processing, files in `ozone`, `pm2.5`, and `so2` folder contain data for New York State, others contain data across the US.

### Ozone, 15 years, New York state

```{r}
full_ozone = 
  tibble(
    files = list.files("data/daily_parameter/ozone/"),
    path = str_c("data/daily_parameter/ozone/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

### CO, 15 years, New York State

```{r}
full_co = 
  tibble(
    files = list.files("data/daily_parameter/co/"),
    path = str_c("data/daily_parameter/co/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

### NO2, 15 years, New York State

```{r}
full_no2 = 
  tibble(
    files = list.files("data/daily_parameter/no2/"),
    path = str_c("data/daily_parameter/no2/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

### SO2, 15 years, New York State

```{r}
full_so2 = 
  tibble(
    files = list.files("data/daily_parameter/so2/"),
    path = str_c("data/daily_parameter/so2/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

### PM2.5, 15 years, New York State

```{r}
full_pm2_5 = 
  tibble(
    files = list.files("data/daily_parameter/pm2.5/"),
    path = str_c("data/daily_parameter/pm2.5/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```

### PM10, 15 years, New York State

```{r}
full_pm10 = 
  tibble(
    files = list.files("data/daily_parameter/pm10/"),
    path = str_c("data/daily_parameter/pm10/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  janitor::clean_names() %>% 
  filter(state_name == "New York") %>% 
  select(state = state_name, city = city_name, county = county_name, site = local_site_name, date = date_local, parameter_name, max = x1st_max_value, mean = arithmetic_mean, units_of_measure, observation_count, latitude, longitude)
```


