---
title: "Asthma and heart disease in NY state - from BRFSS"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)
library(dplyr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

### Data Loading

```{r}
brfss_air_df = 
  read_csv("./data/brfss_with_air.csv") %>% 
  drop_na(county) %>% 
  mutate(
    county_code = as.character(county_code)
  )
```

#### Asthma

##### Figure 1: Trend of the rate for people who have asthma history in NY state among 10 years

```{r}
asthma_history_df = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(year) %>% 
  count(
    year, asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n),
    asthma_status = as.factor(asthma_status)
  ) %>% 
  drop_na(asthma_status)

asthma_history_graph = 
  asthma_history_df %>% 
  filter(asthma_status != "3") %>% 
  group_by(asthma_status) %>%
  ggplot(aes(x = year, y = percent, group = asthma_status, color = asthma_status)) +
  geom_smooth()+
  geom_point()+
  scale_fill_viridis(option = "viridis")+
  labs(
    title = "Trend of the rate for people who have asthma history in NY state among 10 years",
    x = "Year",
    y = "Percent of people who have asthma",
    color = "Asthma Status"
  ) +
  scale_x_continuous(breaks = 2003:2012 )

asthma_history_graph
```

##### Figure 2: Map of the rate for people who currently have asthma among 10 years

```{r}
asthma_now_county_df = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df, values = "1") +
  scale_fill_continuous(
    low = "white", high = "Blue", name = "Rate for people who currently have asthma", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### Figure 3: Trend of average doctor visits for urgent treatment of worsening asthma symptoms in counties among 10 years

```{r}
asthma_visit_df = 
  brfss_air_df %>% 
  drop_na(asthma_visit) %>% 
  filter(asthma_visit != 0) %>% 
  group_by(state_code.x, county_code,county) %>% 
  summarize(
    mean_asthma_visit = mean(asthma_visit),
    max_asthma_visit = max(asthma_visit),
    min_asthma_visit = min(asthma_visit)
  ) 

asthma_visit_graph = 
  asthma_visit_df %>% 
  mutate(
    county = fct_reorder(county, desc(mean_asthma_visit))
  ) %>% 
  ggplot(aes(x = county, y = mean_asthma_visit)) +
  geom_point()+
  geom_errorbar(mapping = aes(ymin = min_asthma_visit, ymax = max_asthma_visit )) +
  labs( x = "County",  y = "Average Doctor Visits for Asthma", title = "Average doctor visits for urgent treatment of worsening asthma symptoms among 10 years") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

asthma_visit_graph
```

#### Heart disease

##### Figure 4: Trend of the rate for people who have any heart disease in NY state among 10 years

```{r}
hd_history_df = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code),
    heart_disease = ifelse(coronary_heart_disease == "1" | heart_attack == "1" | stroke == "1", "1", "0")
  ) %>% 
  group_by(year) %>% 
  count(
    year, heart_disease
  ) %>% 
  mutate(
    percent = n/sum(n),
    heart_disease = as.factor(heart_disease)
  ) %>% 
  drop_na(heart_disease)

hd_history_graph = 
 hd_history_df %>% 
  filter(heart_disease == "1") %>% 
  ggplot(aes(x = year, y = percent)) +
  geom_smooth()+
  geom_point()+
  scale_fill_viridis(option = "viridis")+
  labs(
    title = "Trend of the rate for people who have any heart disease in NY state among 10 years",
    x = "Year",
    y = "Percent of people who have heart disease"
  ) +
  scale_x_continuous(breaks = 2003:2012 )

hd_history_graph
```

##### Figure 5: Map of the rate for people who have heart disease among 10 years

```{r}
hd_county_df = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code),
    heart_disease = ifelse(coronary_heart_disease == "1" | heart_attack == "1" | stroke == "1", "1", "0")
  ) %>% 
  group_by(state_code.x, county_code,county,fips) %>% 
  count(
    year, heart_disease
  ) %>% 
  mutate(
    percent = n/sum(n),
    heart_disease = as.factor(heart_disease)
  ) %>% 
  drop_na(heart_disease)

hd_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = hd_county_df, values = "percent") +
  scale_fill_continuous(
    low = "white", high = "Green", name = "Rate of people who have heart disease", label = scales::comma, limits = c(0,0.8)
  ) + 
  theme(legend.position = "right")

hd_county_plot_map 
```

