---
title: "Type of Pollutants in each counties"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```


```{r, include=FALSE}
air_daily_df = 
  read_csv("./data/air_daily.csv") %>% 
  separate(col = date, into = c('year','month','day'), sep = "-" , convert = TRUE) %>% 
  drop_na(aqi) %>% 
  drop_na(county)
```

```{r, include=FALSE}
aqi_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, aqi) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    aqi_mean = mean(aqi)
  )

air_county_df = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

air_quality_day_df = 
  air_daily_df %>% 
  group_by(state_code, county_code,county) %>% 
  mutate(
    aqi_status = case_when(
      category %in% c("Good", "Moderate") ~ "Healthy",
      category %in% c("Unhealthy for Sensitive Groups", "Unhealthy", "Very Unhealthy") ~ "Unhealthy"
    )
  )
```

#### Figure 5: Defining pollutants distribution among 10 years in NY State

```{r}
pollutants_graph = 
  air_quality_day_df %>% 
  group_by(defining_parameter) %>% 
  summarize(
    pollutants = n()
  ) %>%
  mutate(
    defining_parameter = fct_reorder(defining_parameter, pollutants)
    ) %>% 
  ggplot(aes(x = defining_parameter, y = pollutants, fill = pollutants)) +
  geom_col() +
  labs(
    title = "Defining pollutants distribution among 10 years in NY State",
    x = "Defining pollutants",
    y = "days"
  ) +
  scale_fill_viridis(option = "turbo")

pollutants_graph
```

* Defining pollutants is the most important pollutants in one day in a place. If there is no other pollutants that exceeds the control line, the defining pollutants is `Ozone`. Based on the graph, we can know that except for `Ozone`, `PM2.5` and `NO2` are pollutants that affect NY state in past 10 years.

#### Figure 6: Defining pollutants distribution in unhealthy air quality days among 10 years in NY State

```{r}
Unhealthy_pollutants_graph = 
  air_quality_day_df %>% 
  filter(aqi_status == "Unhealthy") %>% 
  group_by(defining_parameter) %>% 
  summarize(
    pollutants = n()
  ) %>%
  mutate(
    defining_parameter = fct_reorder(defining_parameter, pollutants)
    ) %>% 
  ggplot(aes(x = defining_parameter, y = pollutants, fill = pollutants)) +
  geom_col() +
  labs(
    title = "Defining pollutants distribution in unhealthy air quality days among 10 years in NY State",
    x = "Defining pollutants",
    y = "Unhealthy air quality days"
  ) +
  scale_fill_viridis(option = "turbo")

Unhealthy_pollutants_graph
```

* There are nearly 3000 unhealthy days in total for counties in past 10 years.
* Based on the graph, we can know that `Ozone`and `PM2.5` are main defining pollutants during unhealthy air days. 

#### Ozone

##### Figure 7: Mean Ozone by county in NY state, 2003-2012, by counties

```{r}
ozone_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, mean_ozone) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    ozone_mean = mean(mean_ozone)
  ) %>% 
  drop_na(ozone_mean)

ozone_graph =
  ozone_year_df %>% 
  group_by(county) %>% 
  ggplot(aes(x = year, y = ozone_mean, color = county)) +
  geom_point(alpha=.3) +
  geom_line() +
  labs(
    title = "Mean Ozone by county in NY state, 2003-2012, by counties",
    x = "Year",
    y = "Mean Ozone"
  )+
  scale_x_continuous(breaks = 2003:2012 )+
  scale_color_viridis(
    name = "County", 
    discrete = TRUE
  )

ozone_graph
```

* The United States Environmental Protection Agency test ground-level ozone. Thus, when `ozone` is increasing, air quality is worse. Based on the graph, it can be sen that `ozone` Concentration is increasing among 10 years.
* We can also find that some counties are with a lower `ozone` concentration among 10 years, such as `New York`, `Bronx` and `Queens`. However, some counties are with higher `ozone` concentration, for example, `Tompkins`, `Rensselaer` and `Chautauqua`.

#### Figure 8: Ozone by counties on map

```{r}
ozone_county_df = 
  ozone_year_df %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    ozone_all = mean(ozone_mean),
    max = max(ozone_mean),
    min = min(ozone_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  ) 

ozone_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = ozone_county_df, values = "ozone_all", labels = TRUE, label_color = "white") +
  scale_fill_continuous(
    low = "white", high = "Blue", name = "Air Quality Index", label = scales::comma, limits = c(0,0.04)
  ) +   
  labs(
    title = "Ozone by counties on map"
  )+
  theme(legend.position = "right")

ozone_county_plot_map
```

* Maps can help us directly view the `ozone` concentration in counties. This map is based on mean ozone concentration among 10 years. According to the map, we can see that `Essex`, `Ulster` and `Chautauqua` are with the highest ozone concentration among 10 years. 

#### PM2.5

##### Figure 9: Days for PM2.5 presence by counties within 10 years

```{r}
pm2_5_county_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, mean_pm2_5) %>% 
  drop_na(mean_pm2_5) %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    pm2_5 = n(),
    pm2_5_all = mean(mean_pm2_5),
    max = max(mean_pm2_5),
    min = min(mean_pm2_5)
  ) %>% 
  ungroup() %>% 
  arrange(pm2_5) %>% 
  mutate(
    county = fct_reorder(county, pm2_5)
  )
  
pm2_5_county_graph = 
  ggplot(pm2_5_county_df, aes(y = county, x = pm2_5, fill = pm2_5)) +
  geom_col() +
  labs(
    title = "Days for PM2.5 presence by counties within 10 years",
    x = "Days when PM2.5 can be detected",
    y = "County"
  ) +
  scale_fill_viridis(option = "turbo")
  
pm2_5_county_graph
```

* `PM2.5` is a harmful pollutant that will affect human health. According to the result, `Queens`, `Bronx`, `Steuben`, `Albany` and `Erie` are counties with most PM2.5, while `St. Lawrence`, `Oneida`, `Schenectady`, `Dutchess` and `Broome` are hard to detect PM2.5.

##### Figure 10: PM2.5 distribution by counties within 10 years

```{r}
pm2_5_range_df = 
  pm2_5_county_df %>% 
  mutate(
    county = fct_reorder(county, pm2_5_all)
  )

pm2_5_county_range_graph =   
  pm2_5_range_df %>% 
  group_by(county) %>% 
  ggplot(aes(x = county, y = pm2_5_all)) +
  geom_point()+
  geom_errorbar(mapping = aes(ymin = min, ymax = max)) +
  labs( x = "County",  y = "PM2.5", title = "PM2.5 distribution by counties within 10 years") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

pm2_5_county_range_graph
```

* According to the error bar graph, we can know that there isn't much difference for mean PM2.5 concentration among different counties in 10 years, but we can find that the ranges for PM2.5 concentration are different. `Queens` is the county with the highest PM2.5 observation.
