---
title: "Asthma and air quality"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)
library(dplyr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```


```{r, include=FALSE}
brfss_air_df = 
  read_csv("./data/brfss_with_air.csv") %>% 
  drop_na(county) %>% 
  mutate(
    county_code = as.character(county_code)
  )

brfss_air2_df = 
  read_csv("./data/brfss_with_air2.csv") %>% 
  drop_na(county) %>% 
  mutate(
    county_code = as.character(county_code)
  )
```


#### Figure 1: Association between air quality and Asthma emergency visit

* In order to find the association between air quality and asthma emergency visit times, we use several different predictors and plot graphs.
* Based on the graphs, we can find that `air quality index` and `ozone` are with no relationships with asthma emergency visit times; `CO`, `NO`, `SO2` and `PM2.5` are with positive relationships with asthma emergency visit times. Thus, we can use those predictors for our modeling.

##### air quality index 

```{r}
asthma_air_graph1 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_aqi_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph1
```

##### Ozone

```{r}
asthma_air_graph2 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_ozone_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph2
```

##### CO

```{r}
asthma_air_graph3 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_co_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph3
```

##### NO2

```{r}
asthma_air_graph4 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_no2_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph4
```

##### SO2

```{r}
asthma_air_graph5 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_so2_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph5
```

##### PM2.5

```{r}
asthma_air_graph6 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_pm2_5_month, y = asthma_emergency)) +
  geom_point()+
  geom_smooth()

asthma_air_graph6
```

#### Figure 2: Association between air quality and Asthma doctor visit

* In order to find the association between air quality and asthma doctor visit times, we use several different predictors and plot graphs.
* Based on the graphs, we can find that all predictors are with no relationships with asthma doctor visit times.

##### air quality index

```{r}
asthma_air_graph1 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_aqi_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph1
```

##### Ozone

```{r}
asthma_air_graph2 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_ozone_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph2
```

##### CO

```{r}
asthma_air_graph3 = 
  brfss_air2_df %>%
  ggplot(aes(x = mean_co_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph3
```

##### NO2

```{r}
asthma_air_graph4 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_no2_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph4
```

##### SO2

```{r}
asthma_air_graph5 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_so2_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph5
```

##### PM2.5

```{r}
asthma_air_graph6 = 
  brfss_air2_df %>% 
  ggplot(aes(x = mean_pm2_5_month, y = asthma_visit)) +
  geom_point()+
  geom_smooth()

asthma_air_graph6
```

