---
title: "Air Quality Index across year in NY state"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```


```{r, include=FALSE}
air_daily_df = 
  read_csv("./data/air_daily.csv") %>% 
  separate(col = date, into = c('year','month','day'), sep = "-" , convert = TRUE) %>% 
  drop_na(aqi) %>% 
  drop_na(county)
```


```{r, include=FALSE}
aqi_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, aqi) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    aqi_mean = mean(aqi)
  )
```


#### Figure 1: Air Quality Index by counties in NY State, 2003-2012

```{r}
aqi_state_graph =
  aqi_year_df %>% 
  group_by(county) %>% 
  ggplot(aes(x = year, y = aqi_mean, color = county)) +
  geom_point(alpha=.3) +
  geom_line() +
  labs(
    title = "Air Quality Index by county in NY state, 2003-2012",
    x = "Year",
    y = "Air Quality Index"
  )+
  scale_x_continuous(breaks = 2003:2012 )+
  scale_color_viridis(
    name = "County", 
    discrete = TRUE
  )

aqi_state_graph
```

* According to the graph, we can find that some counties are with a high air quality among 10 years, reaching 50-60(Moderate), for example, `New York`, `Bronx`, `Erie` and `Queens`. However, some counties are with lower air quality index, as lower as 10-30, for example, `Franklin`, `Columbia`,`St. Lawrence` and `Kings`. 
* It can be seen that air quality in most of counties are decreasing among 10 years, which means that air quality in NY state is getting better. 

#### Figure 2: Mean AQI for different counties in NY State, 2003-2012

```{r}
aqi_county_graph = 
  aqi_year_df %>% 
  group_by(county) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(county = fct_reorder(county, aqi_all)) %>% 
  ggplot(aes(x = county, y = aqi_all)) +
  geom_point()+
  geom_errorbar(mapping = aes(ymin = min, ymax = max)) +
  labs( x = "County",  y = "Air Quality Index", title = "Mean AQI for different counties in NY State, 2003-2012") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

aqi_county_graph
```

* This graph is based on mean air quality among 10 years in different counties. According to this graph, we can directly find the top 5 counties with worst air quality(`Bronx`, `New York`, `Queens`, `Erie` and `Richmond`) and top 5 counties with best air quality(`Columbia`, `Franklin`, `St. Lawrence`, `Schenectady` and `Dutchess`)

#### Figure 3: AQI by counties in NY State on map

```{r}
air_county_df = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df, values = "aqi_all", labels = TRUE, label_color = "white") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "Air Quality Index", label = scales::comma, limits = c(0,60)
  ) +   
  labs(
    title = "AQI by counties in NY State on mapp"
  )+
  theme(legend.position = "right")

county_plot_map
```

* Maps can help us directly view the air quality in counties. This map is based on mean air quality index among 10 years. 


#### Figure 4: Unhealthy air quality days by counties in NY State, 2003-2012

```{r}
air_quality_day_df = 
  air_daily_df %>% 
  group_by(state_code, county_code,county) %>% 
  mutate(
    aqi_status = case_when(
      category %in% c("Good", "Moderate") ~ "Healthy",
      category %in% c("Unhealthy for Sensitive Groups", "Unhealthy", "Very Unhealthy") ~ "Unhealthy"
    )
  ) 

Unhealthy_air_graph = 
  air_quality_day_df %>% 
  filter(aqi_status == "Unhealthy") %>% 
  group_by(county) %>% 
  summarize(
    unhealthy_days = n()
  ) %>% 
  mutate(
    county = fct_reorder(county, unhealthy_days)
    ) %>% 
  ggplot(aes(y = county, x = unhealthy_days, fill = unhealthy_days)) +
  geom_col() +
  labs(
    title = "Unhealthy air quality days by counties in NY State, 2003-2012",
    x = "Unhealthy air quality days",
    y = "County"
  ) +
  scale_fill_viridis(option = "turbo")

Unhealthy_air_graph 
```

* According to the classification, the air is defined as unhealthy when air quality index is higher than 100. Based on the graph, we can know that top 5 counties with the most unhealthy air quality days are `Suffolk`, `Chautauqua`, `Richmond`, `Erie` and `Westchester`. 