---
title: "Exploratory Analysis"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```


### Air Quality Index across year in NY state

#### Load the data

```{r}
air_daily_df = 
  read_csv("./data/air_daily.csv") %>% 
  separate(col = date, into = c('year','month','day'), sep = "-" , convert = TRUE) %>% 
  drop_na(aqi) %>% 
  drop_na(county)
```

#### Calculate the mean aqi for each county each year

```{r}
aqi_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, aqi) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    aqi_mean = mean(aqi)
  )
```


#### Figure 1: Aqi for different county from 2003-2012

```{r}
aqi_state_graph =
  aqi_year_df %>% 
  group_by(county) %>% 
  ggplot(aes(x = year, y = aqi_mean, color = county)) +
  geom_point(alpha=.3) +
  geom_line() +
  labs(
    title = "Air Quality Index by county in NY state, 2003-2012",
    x = "Year",
    y = "Air Quality Index"
  )+
  scale_x_continuous(breaks = 2003:2012 )+
  scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

aqi_state_graph
```


##### Figure 2: Mean AQI for different county from 2003-2012 

```{r}
aqi_county_graph = 
  aqi_year_df %>% 
  group_by(county) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(county = fct_reorder(county, aqi_all)) %>% 
  ggplot(aes(x = county, y = aqi_all)) +
  geom_point()+
  geom_errorbar(mapping = aes(ymin = min, ymax = max)) +
  labs( x = "County",  y = "Air Quality Index", title = "Mean AQI for different county from 2003-2012") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

aqi_county_graph
```


#### Figure 3: Map for Aqi in NY state

```{r}
air_county_df = 
  aqi_year_df %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    aqi_all = mean(aqi_mean),
    max = max(aqi_mean),
    min = min(aqi_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = air_county_df, values = "aqi_all") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "Air Quality Index", label = scales::comma, limits = c(0,60)
  ) + 
  theme(legend.position = "right")

county_plot_map
```

#### Figure 4: Unhealthy air quality days in counties among 10 years

```{r}
air_quality_day_df = 
  air_daily_df %>% 
  group_by(state_code, county_code,county) %>% 
  mutate(
    aqi_status = case_when(
      category %in% c("Good", "Moderate") ~ "Healthy",
      category %in% c("Unhealthy for Sensitive Groups", "Unhealthy", "Very Unhealthy") ~ "Unhealthy"
    )
  ) 

Unhealthy_air_graph = 
  air_quality_day_df %>% 
  filter(aqi_status == "Unhealthy") %>% 
  group_by(county) %>% 
  summarize(
    unhealthy_days = n()
  ) %>% 
  mutate(
    county = fct_reorder(county, unhealthy_days)
    ) %>% 
  ggplot(aes(y = county, x = unhealthy_days, fill = unhealthy_days)) +
  geom_col() +
  labs(
    title = "Unhealthy air quality days in counties among 10 years",
    x = "Unhealthy air quality days",
    y = "County"
  ) +
  scale_fill_viridis(option = "turbo")

Unhealthy_air_graph 
```

### Type of Pollutants in each counties

#### Figure 5: Defining pollutants distribution

```{r}
pollutants_graph = 
  air_quality_day_df %>% 
  group_by(defining_parameter) %>% 
  summarize(
    pollutants = n()
  ) %>%
  mutate(
    defining_parameter = fct_reorder(defining_parameter, pollutants)
    ) %>% 
  ggplot(aes(x = defining_parameter, y = pollutants, fill = pollutants)) +
  geom_col() +
  labs(
    title = "Defining pollutants distribution among 10 years in NY state",
    x = "Defining pollutants",
    y = "days"
  ) +
  scale_fill_viridis(option = "turbo")

pollutants_graph
```

#### Figure 6: Defining pollutants distribution in unhealthy air quality days

```{r}
Unhealthy_pollutants_graph = 
  air_quality_day_df %>% 
  filter(aqi_status == "Unhealthy") %>% 
  group_by(defining_parameter) %>% 
  summarize(
    pollutants = n()
  ) %>%
  mutate(
    defining_parameter = fct_reorder(defining_parameter, pollutants)
    ) %>% 
  ggplot(aes(x = defining_parameter, y = pollutants, fill = pollutants)) +
  geom_col() +
  labs(
    title = "Defining pollutants distribution in unhealthy air quality days among 10 years in NY state",
    x = "Defining pollutants",
    y = "Unhealthy air quality days"
  ) +
  scale_fill_viridis(option = "turbo")

Unhealthy_pollutants_graph
```

#### Ozone

##### Figure 7: Defining pollutants distribution in unhealthy air quality days

```{r}
ozone_year_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, mean_ozone) %>% 
  group_by(state_code, county_code,county,year) %>% 
  summarize(
    ozone_mean = mean(mean_ozone)
  ) %>% 
  drop_na(ozone_mean)

ozone_graph =
  ozone_year_df %>% 
  group_by(county) %>% 
  ggplot(aes(x = year, y = ozone_mean, color = county)) +
  geom_point(alpha=.3) +
  geom_line() +
  labs(
    title = "Mean Ozone by county in NY state, 2003-2012",
    x = "Year",
    y = "Mean Ozone"
  )+
  scale_x_continuous(breaks = 2003:2012 )+
  scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ozone_graph
```

#### Figure 8: Map for ozone in NY state

```{r}
ozone_county_df = 
  ozone_year_df %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    ozone_all = mean(ozone_mean),
    max = max(ozone_mean),
    min = min(ozone_mean)
  ) %>% 
  mutate(
    fips = str_c(state_code,county_code)
  )

ozone_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = ozone_county_df, values = "ozone_all") +
  scale_fill_continuous(
    low = "white", high = "Red", name = "Air Quality Index", label = scales::comma, limits = c(0,0.04)
  ) + 
  theme(legend.position = "right")

ozone_county_plot_map
```

#### PM2.5

##### Figure 9: Days when PM2.5 can be detected for each counties 

```{r}
pm2_5_county_df = 
  air_daily_df %>% 
  select(state_code, county_code, state, county, year, mean_pm2_5) %>% 
  drop_na(mean_pm2_5) %>% 
  group_by(state_code, county_code,county) %>% 
  summarize(
    pm2_5 = n(),
    pm2_5_all = mean(mean_pm2_5),
    max = max(mean_pm2_5),
    min = min(mean_pm2_5)
  ) 
  
pm2_5_county_graph = 
  pm2_5_county_df %>% 
  group_by(county) %>% 
  mutate(county = fct_reorder(county, pm2_5)) %>%
  ggplot(aes(y = county, x = pm2_5, fill = pm2_5)) +
  geom_col() +
  labs(
    title = "Days when PM2.5 can be detected for each counties",
    x = "Days when PM2.5 can be detected",
    y = "County"
  ) +
  scale_fill_viridis(option = "turbo")
  
  pm2_5_county_graph
```

##### Figure 10: PM2.5 range for each counties among 10 years

```{r}
pm2_5_county_range_graph =   
  pm2_5_county_df %>% 
  group_by(county) %>% 
  mutate(county = fct_reorder(county, pm2_5_all)) %>% 
  ggplot(aes(x = county, y = pm2_5_all)) +
  geom_point()+
  geom_errorbar(mapping = aes(ymin = min, ymax = max)) +
  labs( x = "County",  y = "PM2.5", title = "PM2.5 range for each counties among 10 years") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

pm2_5_county_range_graph
```


#### BRFSS

```{r}
brfss_air_df = 
  read_csv("./data/brfss_with_air.csv") %>% 
  drop_na(county)
```

#### BRFSS & Air quality
