---
title: "Asthma and heart disease in NY state"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(patchwork)
library(usmap)
library(dplyr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

```{r, include=FALSE}
brfss_air_df = 
  read_csv("./data/brfss_with_air.csv") %>% 
  drop_na(county) %>% 
  mutate(
    county_code = as.character(county_code)
  )
```

* Considering geological factor as an important predictor, we calculated percentage of asthma cases based on their locations with respect to counties in NY state and created maps with their asthma rate based on colors among counties in NY state.If the color is grey, that means there was no respondents in those counties at that year.

#### Asthma Map{.tabset}

According to the graphs, we can know that:
* In 2003, `Chautauqua` county is with the most asthma patients. After that, the percentage decreased until 2009. 
* After 2008, asthma patients in `Steuben` and `Chemung` were increasing.
* Counties in north of New York, including `Jefferson`, `St. Lawrence`, `Franklin` and `Essex` are also with higher percentage of asthma patients.
* Urban counties, including `Bronx`, `Queens` and `Erie`, are also with a higher percentage of asthma patients.

##### 2003

```{r}
asthma_now_county_df1 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2003) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df1, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2004

```{r}
asthma_now_county_df2 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2004) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df2, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2005

```{r}
asthma_now_county_df3 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2005) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df3, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2006

```{r}
asthma_now_county_df4 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2006) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df4, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2007

```{r}
asthma_now_county_df5 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2007) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df5, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2008

```{r}
asthma_now_county_df6 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2008) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df6, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2009

```{r}
asthma_now_county_df7 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2009) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df7, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2010

```{r}
asthma_now_county_df8 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2010) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df8, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2011

```{r}
asthma_now_county_df9 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2011) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df9, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```

##### 2012

```{r}
asthma_now_county_df10 = 
  brfss_air_df %>% 
  mutate(
    fips = str_c(state_code.x,county_code)
  ) %>% 
  group_by(state_code.x, county_code,county,fips, year) %>% 
  filter(year == 2012) %>% 
  count(
    county,asthma_status
  ) %>% 
  mutate(
    percent = n/sum(n)
  ) %>% 
  filter(asthma_status == "1") %>% 
  spread(asthma_status, percent)

asthma_now_county_plot_map = 
  plot_usmap(regions = "county", include = c("NY"), data = asthma_now_county_df10, values = "1", label = TRUE, label_color = "White") +
  scale_fill_continuous(
    low = "white", high = "Orange", name = "Percentage", label = scales::comma, limits = c(0.05,0.17)
  ) + 
  theme(legend.position = "right")

asthma_now_county_plot_map
```