---
title: "Tests_in_website"
output: github_document
---
```{r setup, include=FALSE}
library(plyr)
library(tidyverse)
library(patchwork)
library(viridis)
library(haven)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
asthma_df2 = 
  read_csv("./data/brfss_with_air2.csv") %>%
  mutate(
    #outcome
    asthma = as.numeric(asthma),
    asthma_now = as.numeric(asthma_now),
    #predictors below
    mean_aqi_month = as.numeric(mean_aqi_month),
    mental_health = as.numeric(mental_health),
    physical_health = as.numeric(physical_health),
    county = as.factor(county),
    age = as.factor(age),
    race = as.factor(race),
    smoker = as.factor(smoker),
    income = as.factor(income)
    )
  
```

```{r}
asthma_df2 = asthma_df2 %>%
  mutate(
    aqi_cat = 
         case_when(
           mean_aqi_month <= 50 ~ "Good",
           mean_aqi_month > 50  ~ "Not Good",
           )
      ) 
chisq.test(asthma_df2$aqi_cat, asthma_df2$asthma_now, correct=FALSE)
```
#	Pearson's Chi-squared test

#data:  asthma_df2$aqi_cat and asthma_df2$asthma_now
#X-squared = 2.0033, df = 1, p-value = 0.157


```{r}
asthma_df2_ttest = asthma_df2 %>%
  mutate(
    aqi_cat = 
         case_when(
           mean_aqi_month <= 50 ~ "Good",
           mean_aqi_month > 50  ~ "Not_Good",
           )
      ) %>%
  pivot_wider(
    names_from = "aqi_cat",
    values_from = "asthma_emergency"
  )

t.test(asthma_df2_ttest$Good, asthma_df2_ttest$Not_Good)
```
#	Welch Two Sample t-test

#data:  asthma_df2_ttest$Good and asthma_df2_ttest$Not_Good
#t = -2.5689, df = 421.41, p-value = 0.01055
#alternative hypothesis: true difference in means is not equal to 0
#95 percent confidence interval:
# -1.3715692 -0.1824764
#sample estimates:
#mean of x mean of y 
# 0.718845  1.495868 


#Chi-square test asthma & heart disease
```{r}
asthma_df2_chi = asthma_df2 %>%
  mutate(
    heart_disease = ifelse(coronary_heart_disease == "1" | heart_attack == "1" | stroke == "1", "1", "0")
  ) 
chisq.test(asthma_df2_chi$heart_disease, asthma_df2_chi$asthma_now, correct=FALSE)
```
#Pearson's Chi-squared test

#data:  asthma_df2_chi$heart_disease and asthma_df2_chi$asthma_now
#X-squared = 54.976, df = 1, p-value = 1.22e-13

```{r}
asthma_now_1 =
  asthma_df2 %>%
  drop_na(asthma_now) %>%
  group_by(county) %>%
  filter(asthma_now==1) %>%
  count(asthma_now==1) 

asthma_now_all =
  asthma_df2 %>%
  drop_na(asthma_now) %>%
  group_by(county) %>%
  count() 

data_for_proptest = 
  left_join(asthma_now_1, asthma_now_all, by = "county")

prop.test(data_for_proptest$n.x, data_for_proptest$n.y)
```

#36-sample test for equality of proportions without continuity correction

#data:  data_for_proptest$n.x out of data_for_proptest$n.y
#X-squared = 83.022, df = 35, p-value = 8.893e-06
#alternative hypothesis: two.sided
#sample estimates:
#   prop 1    prop 2    prop 3    prop 4    prop 5    prop 6    prop 7    prop 8    prop 9   prop 10   prop 11   prop 12   prop 13 
#0.6945813 0.7164179 0.6666667 0.8073394 0.7368421 0.6266667 0.7641326 0.7272727 0.7575758 0.8666667 0.8082192 0.7090164 0.4827586 
#  prop 14   prop 15   prop 16   prop 17   prop 18   prop 19   prop 20   prop 21   prop 22   prop 23   prop 24   prop 25   prop 26 
#0.6976744 0.6505747 0.6563758 0.7680000 0.7785714 0.7564576 0.6782178 0.8059701 0.7222222 0.6617916 0.7383178 0.6770186 0.6571429 
#  prop 27   prop 28   prop 29   prop 30   prop 31   prop 32   prop 33   prop 34   prop 35   prop 36 
#0.7482014 0.7415730 0.7462687 0.8333333 0.6954813 0.7000000 0.6967213 0.8620690 0.6851312 0.7271330 
